Contextualizing Climate Ensembles

### 2016-07-20 exam2.tex
found Wickham Survey on boxplots...fills out most of that section

  Solution: THere are a kitchen sinks worth of methods and toolkits for
  visualizing ensembles, but the best ones clearly articulate the
  spatio-temporal relationships in the ensembles and between the variables such
  that the user can definitively parse out what a line or dot is saying about
  the temperature in New Jersey.  
}

@@

Organization:
\begin{table}
data & task & idiom  & AUDIENCE\\

\end{table}

\subsection{Scope}

This paper focuses on the visualization of:
\begin{itemize}
\item climate and weather data - observational and model
\item has both a spaitial and temporal component
\item requires aggregation(ML or stats) before can be visualized with simple viz (line, bar, pie, scatter, heatmap)
\end{itemize}

%%some sorta transition from the generic to the specific tools-and shift this to be more tool oriented

@@

\section{Simple idioms, hard math}

\subsection{Probabilistic aggregations/information gain metrics}
\label{sec:probabilistic}
%%Dump Nir into here-it kinda ties into the band depth stuff that'll show up later
\subsection{markov chain analysis}
\subsection{clustering}

\section{Let's throw Everything in one picture}                                                        
%% Evolution of Single Task/Single Idiom to Higher Order 
\subsection{Scatter, matrix of scatter, scagnostics}                                        
\subsection{Quiver, Flow, Spaghetti, Lasagna}
\subsection{BoxPlots, Functional Box Plots, Countour Box Plots}                                      
A box plot is a really simple measure of uncertainity: there's a line and then two bars represent the uncertainity of that point. %insert image
While this is fairly readable for small datasets, %%\cite{} argues that the upper limit is x/y/z.

>>>>

  Solution: While there are many worth of methods and toolkits for
  visualizing ensembles, the best ones clearly articulate the
  spatio-temporal relationships in the ensembles and between the variables such
  that the user can definitively parse out what a line or dot is saying about
  the temperature in New Jersey.  
}

@@

@@

\section{Visualizing Data & Distributions}

Line, Scatter, Heat...end on the histogram

\section{Visualizing Uncertainty}

While bar graphs and histograms start giving a sense of the variability in
data, they are limited to showing the dispersion of the data only in 1
dimension; namely the measureents themselves. To explore more dimensions,
Tukey introduced the box and whiskers plot \cite{Tukey1970, Tukey1977}. It
takes the distributional information of a histogram and encodes the summary
statistics in a box and outlier information as whiskers coming out of the box,
as seen in %make a figure that shows how a histogram gets encoded as a box and
%whisker.
The advantage of the boxplot is that these boxes can then be strung together,
such that changes in a distribution over time (%show lineplot w/ box & whisker)
or differences over time can be compared. In their survey of the history of the
boxplot \cite{Wickham2011} Wickham and Stryjewski trace the evolution of the
boxplot from a difinitive box to amorphous envelopes.  They start by laying out
the core statistics of a boxplot: a box centered at the median (or mean) and
bound by the upper and lower quartiles, and two whiskers which represent the extremes
(often 1.5 the outer quartiles, sometimes the \alpha bounds). Sometimes there's
also a dot to signify outliers, as seen in \ref{imaginary figure}, but one of
the strengths of the box plot has been in it's flexibility. Many authors have
kept the basic structure, but used different quantile \cite{Hyndman}s or
measures of extremes \cite{Frigge, carter} or used assymetric whiskers \cite{Rousseuw}
or otherwise exploited the structure to incorporate skewness, kurtosis, and
other descriptive distributional statsics \cite{ Aslam, choon, Marmelejo}
%find papers not in Wickham

Inferential information was incorporated through the notched boxplot. \cite{McGill} % find/print this
% paper
In this plot
